---
title: "Første eksempel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Første eksempel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = FALSE
)
```

# DataverkR

DataverkR is a R wrapper for the python library [Dataverk](https://pypi.org/project/dataverk/). It has support for creating and publishing datapackages, with Plotly figures and R dataframes.

### Importing libraries

To create and publish a datapackage you need to import [DataverkR](https://github.com/navikt/dataverkr). Also if you want to add Plotly figures you need to import [Plotly](https://plotly.com/r/).

### Python

You must have Python installed on your machine in order to run this package. The first time you use any python call in R, you will get a warning:

> No non-system installation of Python could be found. Would you like to download and install Miniconda? Miniconda is an open source environment management system for Python. See [miniconda](https://docs.conda.io/en/latest/miniconda.html) for more details.

The warning will only be given one time on each computer and only if Python is not actually installed.

You can install a Python environment in several ways, none of them very good on Windows. The simplest from a desktop installation of R is to download and install it from [Python](https://www.python.org/downloads/) and then follow this excellent guide at [RStudio support](https://support.rstudio.com/hc/en-us/articles/360023654474-Installing-and-Configuring-Python-with-RStudio).

In order to run Python in R, you need the [reticulate](https://cran.r-project.org/package=reticulate) package.

```{r}
install.packages("reticulate")
```

## Creating datapackages

```{r}
library(reticulate)
library(dataverkr)
library(plotly)
```



### Metadata

To create a datapackage you need to supply some metadata in a named list. To see which field are needed see the [datapackages](https://dataplattform.gitbook.io/nada/analyseplattform/dataverk/datapakker) documentation.

```{r}
metadata <- list(title = 'DataverkR example',
                 bucket =  'nav-opendata',
                 store = 'nais', 
                 temporal = reticulate::dict(from = '2020-09-16',
                                             to='2020-09-16'), 
                 type =  'datapackage',
                 format ='datapackage')
```

### Creating the datapackage

To create the datapackage, simply run:

```{r}
dp <- create_dp(metadata)
```

If the metadata is incomplete, an error will occur.

## Adding Plotly Figures and Dataframes

### Dataframes

To add a dataframe:

```{r}
# Creating dataframe
country <- c("Norge","Sverige", "Danmark")
numberOf <- c(200, 300, 500)
df <- data.frame(country, numberOf)

add_resource(dp = dp,
             dataframe = df,
             resource_name = 'some_dataframe',
             resource_description = 'This is a resource example')
```

This dataframe will be available for download as .csv file.

### Plotly figures

To add a Plotly figure:

```{r}
# Creating plotly figure
fig <-  plot_ly(df, x = ~country, y = ~numberOf, type = 'bar')

add_fig(dp = dp,
        fig = fig,
        title = 'Some plotly figure', 
        description = 'This is a plotly figure example')
```

To create a Plotly figure in R, see [Plotly](https://plotly.com/r/).

## Publishing

Finally to publish the datapackage:

```{r}
publish(dp)
```

# For NAV employees

Contact us in [\#data-catalog-intern](https://nav-it.slack.com/archives/CQ9SV9DNE)
