---
title: "Første eksempel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Første eksempel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = FALSE
)
```

# DataverkR

DataverkR is a R wrapper for the python library [Dataverk](https://pypi.org/project/dataverk/). It has support for creating and publishing datapackages, with Plotly figures and R dataframes.

### Importing libraries

To create and publish a datapackage you need to import [DataverkR](https://github.com/navikt/dataverkr). Also if you want to add Plotly figures you need to import [Plotly](https://plotly.com/r/).

### Python

You must have Python installed on your machine in order to run this package. The first time you use any python call in R, you will get a warning:

> No non-system installation of Python could be found. Would you like to download and install Miniconda? Miniconda is an open source environment management system for Python. See [miniconda](https://docs.conda.io/en/latest/miniconda.html) for more details.

The warning will only be given one time on each computer.

You can install a Python environment in several ways, none of them very good on Windows. The simplest from a desktop installation of R is to download and install it from [Python](https://www.python.org/downloads/) and then follow this excellent guide at [RStudio support](https://support.rstudio.com/hc/en-us/articles/360023654474-Installing-and-Configuring-Python-with-RStudio).

But, you probably work at NAV, so any Python way will do. It is also in the internal software kiosk.

It is recommended that you have a virtual Python environment for each and every R project, so as not to let the dependecy hell snakes propagate as it was.

```{python, eval=FALSE}
pip install virtualenv
```

In the terminal, goto your project directory and then create a new virtual environment for Python to live in.

    cd ~/'R prosjekter'/dataverkr

```{python, eval=FALSE}
virtualenv my_env
```

Activate the environment

```{python, eval=FALSE}
source my_env/bin/activate
```

From then on, you can pop into your terminal and install needed Python packages with `pip` without being afraid of messing up EVERYTHING.

To have any use of `dataverkr`, must also have the Python package `dataverk` installed.

> ⚠️ **WARNING** do not pip dataverk without a virtual environment.

Dataverk is a big package with a lot of exotic dependencies. Also, it takes quite a lot of time to complile the packages. Your machine will be on it's knees for a while.

Then, of course, you could use Jupyter onprem or at Google Cloud Platform and not having all these challenges.

```{python, eval=FALSE}
pip install dataverk
```

Then you can load this package and others you might need:

```{r setup}
library(dataverkr)
library(plotly)
```

## Creating datapackages

### Metadata

To create a datapackage you need to supply some metadata in a named list. To see which field are needed see the [datapackages](https://dataplattform.gitbook.io/nada/analyseplattform/dataverk/datapakker) documentation.

```{r}
metadata <- list(title = 'DataverkR example',
                 bucket =  'nav-opendata',
                 store = 'nais', 
                 temporal = reticulate::dict(from = '2020-09-16',
                                             to='2020-09-16'), 
                 type =  'datapackage',
                 format ='datapackage')
```

### Creating the datapackage

To create the datapackage, simply run:

```{r}
dp <- create_dp(metadata)
```

If the metadata is incomplete, an error will occur.

## Adding Plotly Figures and Dataframes

### Dataframes

To add a dataframe:

```{r}
# Creating dataframe
country <- c("Norge","Sverige", "Danmark")
numberOf <- c(200, 300, 500)
df <- data.frame(country, numberOf)

add_resource(dp = dp,
             dataframe = df,
             resource_name = 'some_dataframe',
             resource_description = 'This is a resource example')
```

This dataframe will be available for download as .csv file.

### Plotly figures

To add a Plotly figure:

```{r}
# Creating plotly figure
fig <-  plot_ly(df, x = ~country, y = ~numberOf, type = 'bar')

add_fig(dp = dp,
        fig = fig,
        title = 'Some plotly figure', 
        description = 'This is a plotly figure example')
```

To create a Plotly figure in R, see [Plotly](https://plotly.com/r/).

## Publishing

Finally to publish the datapackage:

```{r}
publish(dp)
```

# For NAV employees

Contact us in [\#data-catalog-intern](https://nav-it.slack.com/archives/CQ9SV9DNE)
